---
title: "Linear Models and Generalized Linear Models in R, using `lm()` and `glm()`"
output: html_notebook
---

First, we are going to discuss how to do a linear regression in R. We are going to focus on the use of linear regression to predict a quantitative response. For example, we might want to learn how to predict the median value of a home, our repsonse variable, based on several possible explanatory variables, such as property-tax, distance to employment centres, number of rooms, etc.

### What can a linear model tell us?

1. Is there a relationship between our response and explanatory variables?

2. How strong is the relationship our response and explanatory variables? 

3. Which explanatory variables contribute to our response variable?

4. How accurately can we estimate the effect of each explanatory variable on the response variable?

5. How accurately can we predict future outcomes

6. Is the relationship linear?

7. Is there synergy among the explanatory variables?

### Explore the data

```{r}
library(MASS)
help(Boston)
str(Boston)
```

Let's start by exploring the relationship between the median value of a home (`medv`), our repsonse variable, and distance to employment centres (`dis`), an explanatory variable.

```{r}
library(ggplot2)
qplot(y = medv, x = dis, data = Boston)
```

There looks as though there may be some relationship, so let's try and fit a model to assess this. 

```{r}
medv_vs_dis <- lm()

```

